A thesis draft

### 2018-11-21 res/sections/08-Background.tex
Move section of background
Minor english fixes

\section{Network slicing}
A network slice can be defined as a logical virtual network, tailored for a
given use case, on the top a common network using technologies such as
Software-Defined Networking and Network Function Chaining that allows network
softwarization~\cite{ordonez2017network}. The notion of separated virtual
networks deployed on the same infrastructure is not new, e.g. VPN, but it has
some specifics that make it a novel idea. The concept of network slicing was
first introduced in 2002 in\cite{peterson2003blueprint}, where a network 
overlay that allows services to run continuously and to access the overlay
resources across different slices was illustrated.
Services operating over the overlay management have their own slice instead.
This idea was lately refined in 2016, where the design of network slicing was
based on a three-layers model~\cite{alliance2016description}, as depicted in 
Figure~\ref{chap:background:img:network_slicing}.

The first layer is the \emph{Service Instance Layer} used to represent the
business or user services supported, the second one is the \emph{Network Slice
Instance Layer}. This is a set of network functions that allows to
satisfy certain requirements~\cite{kotulski2017end}. Network functions, in
general, can be both physical, thus a combination of vendor-specific hardware
and/or virtualized. In that case the function is implemented as software and it
is decoupled from the hardware on which it is running on. Each instance must be
isolated, in order to ensure that function running at the top of it will always
provide the same Quality of Service, without degrading because of a change on
another slice~\cite{richart2016resource}. It also allows to provide better
security. In fact, in case of attack or fault of a certain slice, others must
not be affected. Moreover, each slice must implement it own set of security
function to preventing unauthorized access. Furthermore, isolation let to
manage individually each slice~\cite{ordonez2017network}. Network slice
instances are defined by a \emph{Network Slice Blueprint}: a complete
description of the structure and data flows that allow the creation and the
control of the Network Slice Instance during its whole lifecycle. The blueprint
also refers to the resources needed to assemble the network slice and defines
its characteristics (e.g. ultra-low latency, ultra-reliability, value-added
services for enterprises, etc.). The last layer is the \emph{Resource Layer},
containing both physical and logical resources, that involve all physical
assets and resources for computation, storage or transport, including radio
access.

\section{Software-defined networking}
The SDN is an emerging approach to cloud computing and networking technology
capable of supporting the dynamic nature of future network functions and
intelligent applications while lowering operating costs through simplified
hardware, software, and management~\cite{sezer2013we}, that allows to configure
network in a programmatically way, possibly enhancing monitoring and
performances. One of the main principles of SDN architecture is to decouple the
packet forwarding process, \emph{data plane}, from the routing one, the so
called \emph{control plane}. The architecture also contemplate that network
intelligence and state are logically centralized\cite{fundation2012software}.

@@

In the \emph{Application plane} lays \emph{SDN applications} that are programs
whose role is to set the behavior of the SDN controllers 
and manage its hardware resources and data through APIs, often called 
\emph{northbound interfaces} (NBI). In order to do so they can also build an
abstracted view of the network, using controllers information, in order to make
the traffic analysis easier and to simplify the decision making process..
These interfaces provide an additional abstraction layer that high-level 
applications can consume, enabling the possibility to use a variety of 
different vendors' controllers in the same way. Applications consists both of
the two defined layers, one that is responsible to manage the whole logic, and
another one dedicated to the communication part, that in particular uses the
previously defined NBIs.

@@

\emph{SDN controllers}, are logically centralized components that
receive requests from SDN applications and either translate them 
in a fashion that can be understood from the SDN Datapath, or collect 
data as statics and events needed by the application to build an abstract view
of the network. 
Due to the variety of tasks a controller has to accomplish, its logical 
subdivision is achieved with a three-layer architecture, in which every layer 
exposes it's own interface, that can be consequently used by the upper 
layers, up to the \emph{NBI agents}.
NBI agent manages the communications with SDN applications, exposing interfaces
to access network data and to the set of the controller functionalities.
In particular, the middle layer represents the \emph{SDN Controller Logic},
which process the requests coming from the upper layer applications. Finally
requests are communicated to the remaining network components, exploiting the
Control to Data-Plane Interface driver.

@@

\subsubsection{Data plane}
In the lowest plane we can find the \emph{SDN Datapath}, defined as a logical
network device, which exposes the possibility to access and control to its
forwarding and processing capabilities.It is composed by a CDPI agent and a set
of one or  multiple traffic forwarding engines and of a zero or more traffic
processing functions, that are, as a matter of fact, the devices logic.
Datapath can be mapped either to a single physical device or multiple ones and
the implementation of the its logic does not preclude neither the
virtualization or the sharing of physical resources nor the interoperability
with non-SDN networking.

@@

\section{Virtualization}
Cloud computing is nowadays one of the cutting edge Internet technologies, and
it is possible thanks to virtualization, that allows to isolate applications,
optimizing the utilization of hardware resources and providing more flexibility.
Virtualization, in computing, usually refers to the act of creating a virtual
version of a certain asset, such as virtual computer hardware platform,
operating system (OS), storage device, or computer network
resources~\cite{liu2014research}. Conventional virtualization
(hypervisor-based) uses virtual machines: it provides strong isolation and a
complete operating system~\cite{eder2016hypervisor}. In hypervisor
virtualization every pieces of hardware required to running software is
emulated, and it is common that the hypervisor provides interfaces to access
the real hardware, for instance to access physical devices or to communicate
with the network. However all this functionalities comes to a resource
overhead~\cite{scheepers2014virtualization}Container-based virtualization,
instead, could be a more efficient in certain context, because it does not
emulate the entire computer. A container is seen as an isolate process running
on the host, to which the Linux kernel gives a set of capabilities.

\subsection{Network Virtualization}
Virtualization, in the networking context, refers to the act of combining
hardware, software and functionalities into a virtual network. This abstraction
runs on the top of a physical network in a hyper-visor, decoupling the logical
network from the hardware one. Network virtualization allow multiple,
heterogeneous networks to cohabit, increasing the flexibility and manageability.
The idea of network virtualization came from the decoupling of the role of the
traditional ISP into two different entities: who provides the infrastructure
(InPs) and who provides the services(SPs). In that context it is possible to
aggregate resources, even from different InPs, creating a virtual network,
offering an end-to-end service~\cite{chowdhury2009network}. Most known
instances of NV are \emph{VPNs}, dedicated networks that allows the
communication among multiple sites, using
private and secure tunnels, relying on the public network such as Internet.
Other examples of network virtualization are \emph{overlay networks}. They are
logical networks, build on the top of an existing physical network, typically
implemented at the application layer.

\subsection{Network Function Virtualization}
The \emph{Network Function Virtualization} (NFV) concept refers to the use of 
virtualization technologies to virtualize network nodes functions. This
emerging technology focuses on the decoupling of function logic and hardware,
allowing VNFs to run on the top of servers, switches, storage 
devices or on cloud computing infrastructures. In other word the software logic
is separated from the custom platform on which it runs, enabling the use of
commodity hardware~\cite{gray2016network}.
That approach worked well in the past, because of the need to follow rigorous 
standards, providing at the same time high quality, but it led to long product 
cycles and high specialized services that use a combination of software and
hardware resources, typically vendor-specific, and only largely
non-interoperable within other vendor solutions. New Internet scenarios, that 
include IoT and the increasing number of services to spread multimedia data 
(Skype, Netflix, Spotify, \dots), requires more flexible functions and faster 
network deployments. 
This detachment provides greater flexibility to scale VNF based on traffic and
performances and also improve the automation of function deploy and update:
removing dedicated hardware also will reduce costs machines and for
maintenance.

@@

\subsection{VNF challenges}
This new concept open a plethora of challenges~\cite{han2015network}. Regarding
performances, functions running on general purposes servers cloud affect
throughput and latency, moreover the capacity of the virtualized function can
be less than the hardware and software version. To reach comparable levels in
terms of performances efficient algorithms are needed, as a excellent resource
management. Another problems is the VNF management. In case of hardware
functions resources are almost equivalent, at the contrary, virtualized ones
could have costs and values that differs significantly depending on the point
of presence on which are deployed. On the other hand, since VNFs are software
components, can elastically scale up and down, depending on the traffic loads
and user demand. Another management challenge refers to keeping track where the
VNF is running, that is of paramount importance in case of issues. Related to
possible dysfunctions, software function must reach a certain level of
reliability. Generally, purpose-build network equipment can ensure the
five-nines reliability requirement, that is more challenging to follow in that
scenario due to the higher number of possible points of failure, such as during
the scaling operation or VNF migrating from a point of presence to another.
Finally even security must be taken into account. In fact, infrastructure on
which VNFs run can be outsourced to third parties. Furthermore, orchestration
and management components must implement some sort of intrusion prevention and
detection system and the sharing of resources can introduce security threads.

\section{Service Function Chaining}
The Service Function Chaining (SFC) concept was defined 
in~\cite{halpern2015service} and refers to an ordered set of Service Functions
(SF) and the constraint, defined after a classification process, to packets and
flows. SF are the particular functions whose aim is the treatment of packets
received. They can work on different levels of the network layer. These
components can be implemented as virtual element or a combination of a physical
network element and the software running on the top of it. 
There are no limitations on the number of SFs running on the same network
element, so in general it is possible to have one or more functions running on
the same element. SFs can be aware or not that are part of a certain chain, in
the matter case they are called \emph{SFC aware} functions, in the latter one
\emph{SFC unaware}.

@@

\subsection{Architecture principles}
The SFC technology was though to pursue some concept. One of the most important
principles is the \emph{decoupling}. Decoupling in terms on independence on the
infrastructure and the network, so no change must be performed to the
forwarding underlying topology to deploy SFCs, and in terms of separation
between packet handling (by SFs) and forwarding. This decoupling refers even to
architecture and functions: one must be independent from the other. Another
main point is the possibility to dynamically set routes on the incoming traffic
to decide the best treatment based on its typology. The classification can
occur on different point of the chain and can be updated during the packet
delivery. Finally creation, modification and deletion of a certain SFC must has
To impact on the others.

\subsection{Graph view of an SFC}
An SFC can be seen as a graph in which each node is an SF and edges are the
path through which a packet have to pass through. Because of this definitions
there are no limitation on possible cycles that can occur in the chain, in case
that a certain SF have to be reapplied on the same packet, frame or flow. These
scenarios must be adequately managed to avoid endless loops. This enable even
the possibility of branching on a certain node in order to define the treatment
that has to be imposed from that point of the chain. Traffic path in the SFC
graph are managed and described by the \emph{Service Function Forwarder} (SFF)
and \emph{Service Function Path} (SFP) components. The matter is the element of
the architecture responsible of traffic forwarding. The main goals are to
redirect traffic according to the SFC specified for a specific packet flow,
query the classifier if needed and manage traffic coming back from a SF. The
SFP, instead, is an indirection layer between the fully abstract notion of SFC
and the list of SF that the traffic has to traverse. It is used for dictate the
constraint specification on the path of where packets assigned to a certain
service function path must go. There are no limitation on how specific it must
be: it can specify the exact locations, identifying both the SFF and the SF, or
only give some recommendation or, for example, defining only the SFF and
demanding to it the decision on the SF instance.

@@

The \emph{Network Service Header} (NSH), used by SFC environment for
encapsulation is described in~\cite{rfc8300} and consider three main fields as
represent in Figure~\ref{chap:background:img:nshformat}: Base Header, Service
Path Header and Context Header.
\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{nshformat}
  \caption{Network Service Header~\cite{rfc8300}}
  \label{chap:background:img:nshformat}
\end{figure}

\subsubsection{Base header}
The Base Header utility is provide information about the service header and the
payload protocol.

@@

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.55]{5garchitecture}
  \caption{5G Architecture~\cite{alliance20155g}}
  \label{chap:background:img:5garchitecture}
\end{figure}

@@

\subsubsection{Network slicing}
The role of a \emph{network slice} in a 5G architecture is to specifically
handle the Control-plane of a particular service (e.g. smartphones traffic,
autonomous driving, massive IoT), deploying resources in a manner that assure
the required latency, security and reliability. While some very peculiar legacy
services could require specific hardware, the common resources between services
could be shared in a virtualized way, providing auto-scaling capabilities in
services that are under heavy network pressure.

\section{The VIBES project}
This thesis is part of the VIBES project~\cite{vibesesa}, whose main requirement
is to enhance the performance of end-to-end IP based that involves satellite
connections. To reach this goal, the project specifications suggest to exploit
the 5G incoming technology and use the NFV-MANO architecture to perform first
packets elaboration and performance improvement and finally TCP/IP satellite
chunk optimization with the Performance Enhancing Proxy (PEP). The VIBES project
proposed five technical requirements:
\begin{enumerate}
  \item Analysis of the applicability of current and new Internet protocols in
  the proposed VNF-PEP architecture
  \item Implementation of a VNF-PEP prototype
  \item Building of a PoC test platform
  \item VNF-PEP validation and performance tests on 5G use-cases
  \item Demonstration test-bed management
\end{enumerate}

@@

\subsection{VNF-PEP architecture and internet protocols}
VNF-PEP are the softwarization of usual PEP technology. One of the sake of the
project is to research and include this technology in the usual VNF ecosystem to
include without change satellite communications to the SFC. The architecture of
the systems will be based on clusters of containers whose management and
orchestration will follows the ETSI MANO architecture and SDN principles. Thanks
to VNF-PEP will be possible to dynamically take decisions on how to optimize a
certain communication.

>>>>

\section{Virtualization}
Cloud computing is nowadays one of the cutting edge Internet technologies, and
it is possible thanks to virtualization, that allows to isolate applications,
optimizing the utilization of hardware resources and providing more flexibility.
Virtualization, in computing, usually refers to the act of creating a virtual
version of a certain asset, such as virtual computer hardware platform,
operating system (OS), storage device, or computer network
resources~\cite{liu2014research}. Conventional virtualization
(hypervisor-based) uses virtual machines: it provides strong isolation and a
complete operating system~\cite{eder2016hypervisor}. In hypervisor
virtualization every pieces of hardware required to running software is
emulated, and it is common that the hypervisor provides interfaces to access
the real hardware, for instance to access physical devices or to communicate
with the network. However all this functionalities comes to a resource
overhead~\cite{scheepers2014virtualization}. Container-based virtualization,
instead, could be a more efficient in certain context, because it does not
emulate the entire computer. A container is seen as an isolate process running
on the host, to which the Linux kernel gives a set of capabilities.

\subsection{Network Virtualization}
Virtualization, in the networking context, refers to the act of combining
hardware, software and functionalities into a virtual network. This abstraction
runs on the top of a physical network in a hyper-visor, decoupling the logical
network from the hardware one. Network virtualization allow multiple,
heterogeneous networks to cohabit, increasing the flexibility and manageability.
The idea of network virtualization came from the decoupling of the role of the
traditional ISP into two different entities: who provides the infrastructure
(InPs) and who provides the services (SPs). In that context it is possible to
aggregate resources, even from different InPs, creating a virtual network,
offering an end-to-end service~\cite{chowdhury2009network}. Most known instances
of NV are \emph{VPNs}, dedicated networks that allows the communication among
multiple sites, using private and secure tunnels, relying on the public network
such as Internet. Other examples of network virtualization are \emph{overlay
  networks}. They are logical networks, build on the top of an existing physical
network, typically implemented at the application layer.

\section{Software-defined networking}
The SDN is an emerging approach to cloud computing and networking technology
capable of supporting the dynamic nature of future network functions and
intelligent applications while lowering operating costs through simplified
hardware, software, and management~\cite{sezer2013we}, that allows to configure
network in a programmatically way, possibly enhancing monitoring and
performances. One of the main principles of SDN architecture is to decouple the
packet forwarding process, \emph{data plane}, from the routing one, the so
called \emph{control plane}. The architecture also contemplate that network
intelligence and state are logically centralized~\cite{fundation2012software}.

@@

In the \emph{Application plane} lays \emph{SDN applications} that are programs
whose role is to set the behavior of the SDN controllers 
and manage its hardware resources and data through APIs, often called 
\emph{northbound interfaces} (NBI). In order to do so they can also build an
abstracted view of the network, using controllers information, in order to make
the traffic analysis easier and to simplify the decision making process.
These interfaces provide an additional abstraction layer that high-level 
applications can consume, enabling the possibility to use a variety of 
different vendors' controllers in the same way. Applications consists both of
the two defined layers, one that is responsible to manage the whole logic, and
another one dedicated to the communication part, that in particular uses the
previously defined NBIs.

@@

\emph{SDN controllers} are logically centralized components that
receive requests from SDN applications and either translate them 
in a fashion that can be understood from the SDN Datapath, or collect 
data as statics and events needed by the application to build an abstract view
of the network. 
Due to the variety of tasks a controller has to accomplish, its logical 
subdivision is achieved with a three-layer architecture, in which every layer 
exposes its own interface, that can be consequently used by the upper 
layers, up to the \emph{NBI agents}.
NBI agent manages the communications with SDN applications, exposing interfaces
to access network data and to the set of the controller functionalities.
In particular, the middle layer represents the \emph{SDN Controller Logic},
which process the requests coming from the upper layer applications. Finally
requests are communicated to the remaining network components, exploiting the
Control to Data-Plane Interface driver.

@@

\subsubsection{Data plane}
In the lowest plane we can find the \emph{SDN Datapath}, defined as a logical
network device, which exposes the possibility to access and control to its
forwarding and processing capabilities. It is composed by a CDPI agent and a set
of one or multiple traffic forwarding engines and of a zero or more traffic
processing functions that are, as a matter of fact, the devices logic. Datapath
can be mapped either to a single physical device or multiple ones and the
implementation of the its logic does not preclude neither the virtualization or
the sharing of physical resources nor the interoperability with non-SDN
networking.

@@

\section{Network slicing}
A network slice can be defined as a logical virtual network, tailored for a
given use case, on the top a common network using technologies such as
Software-Defined Networking and Network Function Chaining that allows network
softwarization~\cite{ordonez2017network}. The notion of separated virtual
networks deployed on the same infrastructure is not new, e.g. VPN, but it has
some specifics that make it a novel idea. The concept of network slicing was
first introduced in 2002 in~\cite{peterson2003blueprint}, where a network 
overlay that allows services to run continuously and to access the overlay
resources across different slices was illustrated.
Services operating over the overlay management have their own slice instead.
This idea was lately refined in 2016, where the design of network slicing was
based on a three-layers model~\cite{alliance2016description}, as depicted in 
Figure~\ref{chap:background:img:network_slicing}.

The first layer is the \emph{Service Instance Layer} used to represent the
business or user services supported, the second one is the \emph{Network Slice
Instance Layer}. This is a set of network functions that allows to
satisfy certain requirements~\cite{kotulski2017end}. Network functions, in
general, can be both physical, thus a combination of vendor-specific hardware
and/or virtualized. In that case the function is implemented as software and it
is decoupled from the hardware on which it is running on. Each instance must be
isolated, in order to ensure that function running at the top of it will always
provide the same Quality of Service, without degrading because of a change on
another slice~\cite{richart2016resource}. It also allows to provide better
security. In fact, in case of attack or fault of a certain slice, others must
not be affected. Moreover, each slice must implement its own set of security
function to preventing unauthorized access. Furthermore, isolation let to
manage individually each slice~\cite{ordonez2017network}. Network slice
instances are defined by a \emph{Network Slice Blueprint}: a complete
description of the structure and data flows that allow the creation and the
control of the Network Slice Instance during its whole lifecycle. The blueprint
also refers to the resources needed to assemble the network slice and defines
its characteristics (e.g. ultra-low latency, ultra-reliability, value-added
services for enterprises, etc.). The last layer is the \emph{Resource Layer},
containing both physical and logical resources, that involve all physical
assets and resources for computation, storage or transport, including radio
access.

\subsection{Network Function Virtualization}
The \emph{Network Function Virtualization} (NFV) concept refers to the use of
virtualization technologies to virtualize network nodes functions. This emerging
technology focuses on the decoupling of function logic and hardware, allowing
VNFs to run on the top of servers, switches, storage devices or on cloud
computing infrastructures. In other words the software logic is separated from
the custom platform on which it runs, enabling the use of commodity
hardware~\cite{gray2016network}. That approach worked well in the past, because
of the need to follow rigorous standards, providing at the same time high
quality, but it led to long product cycles and high specialized services that
use a combination of software and hardware resources, typically vendor-specific,
and only largely non-interoperable within other vendor solutions. New Internet
scenarios, that include IoT and the increasing number of services to spread
multimedia data (Skype, Netflix, Spotify, \dots), requires more flexible
functions and faster network deployments. This detachment provides greater
flexibility to scale VNF based on traffic and performances and also improve the
automation of function deploy and update: removing dedicated hardware also will
reduce costs machines and for maintenance.

@@

\subsection{VNF challenges}
This new concept open a plethora of challenges~\cite{han2015network}. Regarding
performances, functions running on general purposes servers cloud affect
throughput and latency, moreover the capacity of the virtualized function can be
less than the hardware and software version. To reach comparable levels in terms
of performances efficient algorithms are needed, as a excellent resource
management. Another problems is the VNF management. In case of hardware
functions resources are almost equivalent, at the contrary, virtualized ones
could have costs and values that differs significantly depending on the point of
presence on which are deployed. On the other hand, since VNFs are software
components, they can elastically scale up and down, depending on the traffic
loads and user demand. Another management challenge refers to keeping track
where the VNF is running, that is of paramount importance in case of issues.
Related to possible dysfunctions, software function must reach a certain level
of reliability. Generally, purpose-build network equipment can ensure the
five-nines reliability requirement, that is more challenging to follow in that
scenario due to the higher number of possible points of failure, such as during
the scaling operation or VNF migrating from a point of presence to another.
Finally even security must be taken into account. In fact, infrastructure on
which VNFs run can be outsourced to third parties. Furthermore, orchestration
and management components must implement some sort of intrusion prevention and
detection system and the sharing of resources can introduce security threads.

\section{Service Function Chaining}
The Service Function Chaining (SFC) concept was defined
in~\cite{halpern2015service} and refers to an ordered set of Service Functions
(SF) and the constraint, defined after a classification process, to packets and
flows. SF are the particular functions whose aim is the treatment of packets
received. They can work on different levels of the network layer. These
components can be implemented as virtual elements only or in a combination of a
physical and software network elements. There are no limitations on the number
of SFs running on the same network element, so in general it is possible to have
one or more functions running on the same element. SFs can be aware or not that
are part of a certain chain, in the matter case they are called \emph{SFC aware}
functions, in the latter one \emph{SFC unaware}.

@@

\subsection{Architecture principles}
The SFC technology was though to pursue some concept. One of the most important
principles is the \emph{decoupling}. Decoupling in terms on independence on the
infrastructure and the network, so no change must be performed to the forwarding
underlying topology to deploy SFCs, and in terms of separation between packet
handling (by SFs) and forwarding. This decoupling refers even to architecture
and functions: one must be independent from the other. Another main point is the
possibility to dynamically set routes on the incoming traffic to decide the best
treatment based on its typology. The classification can occur on different point
of the chain and can be updated during the packet delivery. Finally creation,
modification and deletion of a certain SFC must not impact on the others.

\subsection{Graph view of an SFC}
An SFC can be seen as a graph in which each node is an SF and edges are the path
through which a packet have to pass through. Because of this definitions there
are no limitations on possible cycles that can occur in the chain, in case that
a certain SF has to be reapplied on the same packet, frame or flow. These
scenarios must be adequately managed to avoid endless loops. This enable even
the possibility of branching on a certain node in order to define the treatment
that has to be imposed from that point of the chain. Traffic path in the SFC
graph is managed and described by the \emph{Service Function Forwarder} (SFF)
and \emph{Service Function Path} (SFP) components. The latter is the element of
the architecture responsible of traffic forwarding. The main goals are to
redirect traffic according to the SFC specified for a specific packet flow,
query the classifier if needed and manage traffic coming back from a SF. The
SFP, instead, is an indirection layer between the fully abstract notion of SFC
and the list of SF that the traffic has to traverse. It is used for dictate the
constraint specification on the path of where packets assigned to a certain
service function path must go. There are no limitations on how specific it must
be: it can specify the exact locations, identifying both the SFF and the SF, or
only give some recommendation or, for example, defining only the SFF and
demanding to it the decision on the SF instance.

@@

The \emph{Network Service Header} (NSH), used by SFC environment for
encapsulation is described in~\cite{rfc8300} and it consider three main fields
as represented in Figure~\ref{chap:background:img:nshformat}: Base Header,
Service Path Header and Context Header.
\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{nshformat}
  \caption{Network Service Header~\cite{rfc8300}}
  \label{chap:background:img:nshformat}
\end{figure}

\subsubsection{Base header}
The Base Header utility is to provide information about the service header and
the payload protocol.

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.55]{5garchitecture}
  \caption{5G Architecture~\cite{alliance20155g}}
  \label{chap:background:img:5garchitecture}
\end{figure}

@@

\subsubsection{Network slicing}
The role of a \emph{network slice} in a 5G architecture is to specifically
handle the Control-plane of a particular service (e.g. smartphones traffic,
autonomous driving, massive IoT), deploying resources in a manner that assure
the required latency, security and reliability. While some very peculiar and
legacy service could require specific hardware, the common resources between
services could be shared in a virtualized way, providing auto-scaling
capabilities in services that are under heavy network pressure.

\section{The VIBES project}
This thesis started as a part of the VIBES project~\cite{vibesesa}, whose main
requirement is to enhance the performance of end-to-end IP based that involves
satellite connections. To reach this goal, the project specifications suggest to
exploit the 5G incoming technology and use the NFV-MANO architecture to perform
first packets elaboration and performance improvement and finally TCP/IP
satellite chunk optimization with the Performance Enhancing Proxy (PEP). The
VIBES project proposed five technical requirements:
\begin{enumerate}
  \item Analysis of the applicability of current and new Internet protocols in
  the proposed VNF-PEP architecture
  \item Implementation of a VNF-PEP prototype
  \item Building of a PoC test platform
  \item VNF-PEP validation and performance tests on 5G use-cases
  \item Demonstration test-bed management
\end{enumerate}

@@

\subsection{VNF-PEP architecture and internet protocols}
VNF-PEP is the softwarization of the usual PEP technology. One of the goals of
the project is to research and include this technology in the usual VNF
ecosystem to include without change satellite communications to the SFC. The
architecture of the systems will be based on clusters of containers whose
management and orchestration will follows the ETSI MANO architecture and SDN
principles. Thanks to VNF-PEP will be possible to dynamically take decisions on
how to optimize a certain communication.
